---
title: "STA 523 Final Project"
author: "Andy Cooper, Jingyi Zhang, Mingjie Zhao, Ziwei Zhu"
date: "11/28/2018"
output: html_document
runtime: shiny
---

The first part of the markdown file creates helper functions to be used in the shiny app.

```{r message = F, echo = F}
library(tidyverse)
library(shiny)
library(shinydashboard)
library(janitor)
library(tools)
```


```{r}
#Load data prepared by parse_data.R
players_data <- readRDS("players_data.rds")
```


```{r echo = F}
#Function to get roster for given season
get_roster <- function(s){
  return(players_data %>% filter(season == s)) %>% 
    .$name %>% 
    unique() %>% 
    sort()
}
```

```{r}
#Function to plot season performance for given player, season, and statistic
plot_season <- function(p, s, c){
  avg_stat <- mean(players_data %>% 
                     filter(name == p, category == c) %>%
                     .$quantity)
  g <- players_data %>% 
    mutate(win = factor(win, levels = c(0,1), labels = c("loss", "win"))) %>% 
    filter(name == p, season == s, category == c) %>% 
    ggplot(aes(x = date, y = quantity, color = factor(win))) + 
    geom_point(aes(size = (abs(points_duke - points_other)))) + 
    geom_line(color = "black") + 
    geom_hline(yintercept = avg_stat, col = "#4169e1") +
    labs(x = "Date", y = c %>% toTitleCase(), 
         title = "Click on a point on the plot to see the game summary") +
    scale_color_manual(values = c("red", "#4169e1")) + 
    guides(color = guide_legend(title = "Game Outcome"),
           size = guide_legend(title = "Point Deficit"))
  return(g)
}
```


```{r}
#Function to create map with player's hometown
plot_map <- function(s, p){
  states <- map_data("state")
  coords <- players_data %>% 
    filter(season == s, name == p) %>% 
    select(lat, lon) %>% 
    mutate(lon = lon*-1)
  
  g <- ggplot(data = states) + 
    geom_polygon(aes(x = long, y = lat, group = group), fill = "grey") + 
    geom_point(data = coords, color = "blue",
               mapping = aes(x = lon, y = lat),        
               alpha = 0.5, show.legend = FALSE) +
    coord_fixed(1.3) + 
    theme_void()
  return(g)
}
```


Now the actual shiny app is created. The application has 3 inputs, which are dropdown menus to select the season, the player from that season, and the statistic. The outputs are a player image and summary, a line plot, and a modal dialog box created when points on the plot are clicked.

The application is designed to have three main features. The first feature is player-related information. The app outputs the selected player's jersey number, year, high school, height, weight, and position, along with an image of the player. The app also plots a map with a point showing that player's hometown.

The second feature is player performance information. The selected season, player, and statistic are used to create a ggplot of that player's performance over the course of the season.

The third feature is game-related information. The line plot is interactive; a user can click on a point to get a summary of that game in a modal dialog box. The box contains the exact date of the game, the opponent, and final score. It also includes the names and pictures of the top 3 Duke performers in the chosen statistic for that game.

```{r}
#Shiny UI
ui <- fluidPage(
  titlePanel("Duke Men's Basketball Player Analysis App"),
  sidebarLayout(
    sidebarPanel = sidebarPanel(
      
      #Input season
      selectInput(inputId = "select_season", label = h4("Select Season"), 
                  choices = players_data$season %>% 
                    unique() %>% sort() %>% rev()),
      
      #Input player
      selectInput(inputId = "select_player", 
                  label = h4("Select Player From Season"),
                  choices = players_data$name %>% unique() %>% sort()),
      
      #Input statistic
      selectInput(inputId = "select_category", label = h4("Select Statistic"),
                  choices = players_data$category %>% unique() %>%
                    toTitleCase(),
                  selected = 1),
      HTML("<center>"),
      tags$img(src = "https://upload.wikimedia.org/wikipedia/commons/e/e1/Duke_Athletics_logo.svg",
               width = "50%"),
      HTML("</center>")
    ),
    
    mainPanel = mainPanel(
      fluidRow(
        column(6, uiOutput("player_summary")),
        column(3, uiOutput("player_image")),
        column(3, plotOutput("player_hometowns", height = "125px"))
      ),
      fluidRow(
        plotOutput("player_performance", click = "click_game")
      )
    )
  )
)
```


```{r}
#Shiny server
server <- function(input, output, session){
  
  #Reactive list of player info
  state <- reactiveValues(
    player_info = list()
  )
  
  #Update roster based on season input
  observeEvent(input$select_season, {
    same_player <- ifelse(
      (input$select_player %in% get_roster(input$select_season)),
      input$select_player,
      get_roster(input$select_season)[1])
    
    updateSelectizeInput(session = session, inputId = "select_player", 
                         choices = get_roster(input$select_season),
                         selected = same_player)
  })
  
  #Update player info based on user input
  observe({
    state$player_info <- players_data %>% 
      filter(name == input$select_player & season == input$select_season) %>% 
      select(name, number, position, height, 
             weight, year, hometown, picture) %>% 
      unique()
  })
  
  #Output player summary
  output$player_summary <- renderUI({
    tags$p(paste("Jersey:", state$player_info$number[1]),
           HTML("<br/>"),
           paste("Year:", state$player_info$year[1]),
           HTML("<br/>"),
           paste("Position:", state$player_info$position[1]),
           HTML("<br/>"),
           paste("Hometown:", state$player_info$hometown[1]),
           HTML("<br/>"),
           paste("Height:", state$player_info$height[1]),
           HTML("<br/>"),
           paste0("Weight: ", state$player_info$weight[1], "lbs")
    )
  })
  
  #Output player image
  output$player_image <- renderUI({
    tags$img(src = state$player_info$picture[1], height = "125px")
  })
  
  #Output map of player's hometown
  output$player_hometowns <- renderPlot({
    plot_map(input$select_season, input$select_player)
  })
  
  
  #Create line graph
  output$player_performance <- renderPlot({
    plot_season(input$select_player, 
                input$select_season, input$select_category %>% tolower())
  })
  
  
  #Create game summary in modal dialog box
  observeEvent(input$click_game, {
    
    #Get data rows from user's click
    game_data <- nearPoints(players_data, input$click_game, 
                            threshold = 10, maxpoints = 1,
                            addDist = F) %>% 
      select(date, opponent, win, points_duke, points_other, quantity) %>% 
      unique()
    
    #If a game was clicked on, create modal dialog box
    if(nrow(game_data) > 0){
      
      #Get top 3 players in chosen statistic
      top_3 <- players_data %>% 
        filter(date == game_data$date, 
               category == tolower(input$select_category)) %>% 
        arrange(desc(quantity)) %>% 
        select(name, picture, quantity)
      
      #Create modal dialog box
      showModal(modalDialog(title = paste0("Duke vs. ", game_data$opponent,
                                           " (", 
                                           game_data$date %>% 
                                             format("%B  %d, %Y"),
                                           ")"),
                            fluidRow(
                              column(6, HTML("<center>"),
                                     tags$h3(paste("Duke:",
                                                   game_data$points_duke)),
                                     HTML("</center>")),
                              column(6, HTML("<center>"),
                                     tags$h3(paste0(game_data$opponent, ": ",
                                                    game_data$points_other)),
                                     HTML("<center>"))
                            ),
                            fluidRow(
                              HTML("<center>"), 
                              tags$h3(paste("Top 3 Players in",
                                            input$select_category)),
                              HTML("</center>")
                            ),
                            fluidRow(
                              column(4, 
                                     HTML("<center>"),
                                     tags$img(src = top_3$picture[1],
                                              height = "150px"),
                                     tags$br(),
                                     tags$p(top_3$name[1]),
                                     tags$br(),
                                     tags$h4(top_3$quantity[1]),
                                     HTML("</center>")),
                              column(4, 
                                     HTML("<center>"),
                                     tags$img(src = top_3$picture[2],
                                              height = "150px"),
                                     tags$br(),
                                     tags$p(top_3$name[2]),
                                     tags$br(),
                                     tags$h4(top_3$quantity[2]),
                                     HTML("</center>")),
                              column(4, 
                                     HTML("<center>"),
                                     tags$img(src = top_3$picture[3],
                                              height = "150px"),
                                     tags$br(),
                                     tags$p(top_3$name[3]),
                                     tags$br(),
                                     tags$h4(top_3$quantity[3]),
                                     HTML("</center>"))
                            ),
                            easyClose = T))
    }
  })
  
}
```

```{r}
#Run shiny app
shinyApp(ui, server)
```

